### 2014-01-27 notes/notes.tex
Adding notes

\documentclass[12pt]{article}
\usepackage{parskip}
\usepackage[margin=1in]{geometry}
\usepackage{natbib}

\title{Test statistic selection for randomization inference under interference}
\author{Peter Aronow and Jake Bowers}
\date{\today}

\begin{document}
\maketitle

%% 4 pages

 \citep{bowers2013sutva} followed Rosenbaum by recommending effect increasing
 statistic.

 controls Type I error, but power can be very poor for these types of models.
 We explain why and propose a refinement. Multivariate dependence of
 parameters and outcome space. A weighted E-statistic.

Fisher's sharp null hypothesis of no effects combined with his
randomization-based hypothesis testing framework provided a very clear and
clean vision of what it means to do statistical inference \citep[Chap
2]{fisher:1935}. In recent years, a number of scholars have realized that this
clarity also implied great flexibility in regard to the scientific questions
that one might engage: for example, one could use the sharp null itself to
detect interference among units \citep{aronow-general}, and sharp hypotheses
about effects (in contrast to hypotheses about no effects) have been developed
to assess models of interference \citep{bowers2013sutva}, [ include Rosenbaum
2007, Bowers and Hansen 2009, and a bunch of other stuff by Rosenbaum ].

One problem that has become apparent, as scholars have extended Fisher's basic
ideas, is that as the models of effects have become more complex, the power of
the testing procedure itself has become more important. A simple test of the
sharp null can be executed with power that varies as a function of the design
of the study (proportion treated, blocking structure, etc..), characteristics
of the outcome, and the way that a test statistic summarizes the outcome. Yet,
when the models are complex,  \citet{bowers2013sutva} showed that sometimes a
test will have (1) no power to address the model at all (a function, in part
of the model itself --- for example ...) such that all hypothesized parameters
would receive the same implausibility assessment; (2) or might reject all such
hypothesized parameters. Thus, although in theory one may assess sharp
hypotheses about multiparameter settings, in practice one may not learn much
from such tests --- or, in the absence of simulation studies assessing the
model and test statistic, one may be mislead.

In this paper we use a very simple causal model and a simple test statistic to
first show how the combination of model and test statistic is a crucial choice
--- and how more attention ought to be paid to both the model and the test
statistic.

Then we suggest one test statistic that we think should have good properties
when the number of parameters in a model is low (i.e. $k << n$). And we should
this to be the case --- that the robust F-stat from a linear model is a better
way to go than the KS-test and certainly better than any single measure of
distributional difference.

Finally, we highlight the similarities between testing the sharp hypothesis
and the weak hypothesis --- which in turn allows even more flexibility and
perhaps better performing testing procedures.

Then we suggest a few alternative avenues to pursue.

Some models are well powered for some designs and not other designs.

Using same simulation study using same data from PA study. Better than KS
test.

The F-test is model consistent. By design it controls Type I error, but if the
model is correct, it has some optimality criterion. If you had control
variables, for example, then it would be consistent against mean differences.

Under the model, model consistent.

\bibliographystyle{apsr}
\bibliography{/Users/jwbowers/Documents/PROJECTS/BIB/big}

\end{document}