### 2014-10-06 paper/paper.tex

\documentclass[12pt]{article} 
\RequirePackage{snapshot}

\author{Peter Aronow \thanks{Assistant Professor, Dept of Political Science,
    Yale University} \and Jake Bowers \thanks{Associate Professor, Dept of Political
    Science and Statistics, University of Illinois @ Urbana-Champaign
    (\texttt{jwbowers@illinois.edu}). \textit{Acknowledgements:} Part
    of this work funded by NSF Grants SES-0753168 and SES-0753164
    (joint with Ben Hansen, Dept of Statistics, University of
    Michigan).} \and Mark Fredrickson \thanks{Ph.D. Student in Statistics and Political Science, University of Illinois @ Urbana-Champaign.}}

  \date{\today}

\usepackage{../styles/common}

%%% Paper specific packages
\usepackage{setspace}
\usepackage[letterpaper,bottom=1in,top=1in,right=1in,left=1in,includemp=FALSE]{geometry}
\usepackage{microtype} %only useful with pdflatex

%%% Document Structure Formatting
\usepackage[compact,bottomtitles]{titlesec}
%\titleformat{ ⟨command⟩}[⟨shape⟩]{⟨format⟩}{⟨label⟩}{⟨sep⟩}{⟨before⟩}[⟨after⟩]
\titleformat{\section}[hang]{\large\bfseries}{\thesection}{.5em}{\hspace{0in}}[\vspace{-.2\baselineskip}]
\titleformat{\subsection}[hang]{\bfseries}{\thesubsection}{.5em}{\hspace{0in}}[\vspace{-.2\baselineskip}]
\titleformat{\subsubsection}[hang]{\itshape}{\thesubsubsection}{.5em}{\hspace{0in}}[\vspace{-.2\baselineskip}]
\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1ex}{}{\vspace{-.2\baselineskip}}

%\titlespacing*{ ⟨command⟩}{⟨left⟩}{⟨beforesep⟩}{⟨aftersep⟩}[⟨right⟩]
\titlespacing{\section}{0pc}{1.5ex plus .1ex minus .2ex}{.5ex plus .1ex minus .1ex}
\titlespacing{\subsection}{0pc}{1.5ex plus .1ex minus .2ex}{.5ex plus .1ex minus .1ex}
\titlespacing{\subsubsection}{0pc}{1.5ex plus .1ex minus .2ex}{.5ex plus .1ex minus .1ex}

\newif\ifanon

% \ifanon
% \footnote{We are awesome!}
% \else
% \fi

\begin{document}
\VerbatimFootnotes
\input{titlepage}
%\setstretch{1.6} %1.6 or 1.37 which I think is "true double spacing"
%\onehalfspacing
%%I don't think I need to double space footnotes for AJPS. Check this. 
\abovedisplayshortskip=0pt
\belowdisplayshortskip=0pt
\abovedisplayskip=0pt
\belowdisplayskip=0pt
%\tableofcontents

% include stuff here!
% include stuff here!

\input{introduction.tex}

\cleardoublepage
%\appendix % is redefined by hyperref and a mess, do the appendix stuff by hand instead
% problems with pdf bookmarks and appendix caused by resetting section counter.
% solution here: http://www.latex-community.org/forum/viewtopic.php?f=45&t=4950
\begin{singlespacing}
%\renewcommand{\thesection}{Appendix \thesection}
%\setcounter{section}{0}
\newcounter{mycounter}
\setcounter{mycounter}{0}
%\renewcommand{\thesection}{Appendix \Alph{mycounter}}
\let\osection\section
\renewcommand{\section}{\stepcounter{mycounter}\osection}
\renewcommand\thesection{Appendix \Alph{mycounter}}
%\titleformat{\section}[hang]{\large\bfseries}{\thesection}{.5em}{\hspace{0in}}[]
%\input{codeexample}
%%\clearpage
%%\input{generalinterference}

\clearpage
\bibliographystyle{apsr} 
\bibliography{../BIB/big}

% this appendix is the response to reviewers. TODO: remove for released
% version
% \clearpage
% \setlength{\parindent}{0pt}
% \addtolength{\parskip}{.5\baselineskip}
% \pagenumbering{arabic} %\setcounter{page}{1}
% \setcounter{mycounter}{0}
% \setcounter{secnumdepth}{0}
% %\let\osection\section
% %\renewcommand{\section}{\stepcounter{mycounter}\osection}
% \renewcommand\thesection{\arabic{mycounter}}
% \input{response}
% 
\end{singlespacing}
%\bibliography{/Users/jwbowers/Documents/BIB/trunk/big}
\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: