### 2014-10-06 paper/titlepage.tex

\begin{titlepage}
\maketitle

%\SweaveOpts{echo=true}

\begin{abstract}

\end{abstract}

Key Words: Causal effect; Interference; Randomized experiment;
Randomization inference; Fisher's Sharp Null Hypothesis; SUTVA

\end{titlepage}