### 2015-08-04 paper/titlepage.tex
Adding an abstract to titlepage. Now in introduction.tex itself. Later delete this file

\begin{titlepage}

\maketitle

%\SweaveOpts{echo=true}

\begin{abstract} 
\citet{bowers2013sutva} showed that one could use Fisher's randomization-based
hypothesis testing framework to assess counterfactual causal models of
treatment propagation and spillover across social networks.  This research
note improves the statistical infernce presented in \citet{bowers2013sutva}
(henceforth BFP) by substituting a test statistic based on a sum of squared
residuals and incorporating information about the fixed network for the simple
the Kolmogorov-Smirnov test statistic  \citep[\S 5.4]{MylesHollander1999a}
that they used. The point of this short article is to incrementally improve
the use of BFP's ``reasoning about interference'' approach. We do not offer
results about test statistics for multi-parameter causal models on social
networks here, but instead hope that this research note stimulates further,
and deeper, work on test statistics and sharp hypotheses.
\end{abstract}

Key Words: Causal effect; Interference; Randomized experiment;
Randomization inference; Fisher's Sharp Null Hypothesis; SUTVA

\end{titlepage}