### 2015-11-10 response.md
Sketch/structure for response to reviewers

Dear Professor Grimmer,

We really appreciate the constructive and timely feedback from you and the reviewers.

1) As Reviewer 1 suggests, how does the model perform as the assumptions
about spill overs deviate?  How sensitive are the inferences to the
particular assumptions you can make?

from peter I think we reiterate that this is about testing sharp null
hypotheses, and this will be sensitive to mean deviations w.r.t.
exposure. Any test statistic will yield valid inferences, but SSR is
sensitive to "main effects."

2) Reviewer 2 asks you to carry forward your toy example with four units
through the entire analysis.  If this isn't possible, please provide the
simplest case that illuminate the core intuition of your method.

We have done this.

3) Reviewer 2 also asks you to provide intuition about when the KS-test
would outperform the Sum of Squared Residuals.  As part of a simulation or
analysis of a simple case, demonstrating when/why the KS test outperforms
will be useful for gaining intuition about why the SSR is a generally
superior method.

We have added some discussion about the idea that the SSR is particular sensitive to differences in means.

4) In letters we have an emphasis that new methods matter when applied to
real examples.  Can you demonstrate that a case where the SSR will change
the inferences we make when applied to a real world example?

We have added a section in which we replicate the analysis in Coppock's 2014 Journal of Experimental Political Science article "Information Spillovers: Another Look at Experimental Estimates of Legislator Responsiveness".