### 2015-12-19 paper/paper.tex

\documentclass[12pt]{article}
\RequirePackage{snapshot}
\usepackage{../styles/common}

\author{Jake Bowers \thanks{Associate Professor, Dept of Political Science and Statistics, University of Illinois @ Urbana-Champaign (\texttt{jwbowers@illinois.edu}).} \and Mark Fredrickson \thanks{Ph.D. Student in Statistics and Political Science, University of Illinois @ Urbana-Champaign.} \and Peter M. Aronow \thanks{Assistant Professor, Dept of Political Science and Biostatistics, Yale University}}

%%% Paper specific packages
\usepackage{setspace}
\usepackage[letterpaper,bottom=1in,top=1in,right=1in,left=1in,includemp=FALSE]{geometry}
%\usepackage{microtype} %only useful with pdflatex
%\usepackage{fancyvrb} % for verbatim text in footnotes

%%% Document Structure Formatting
\usepackage[compact,bottomtitles]{titlesec}
%\titleformat{ ⟨command⟩}[⟨shape⟩]{⟨format⟩}{⟨label⟩}{⟨sep⟩}{⟨before⟩}[⟨after⟩]
\titleformat{\section}[hang]{\large\bfseries}{\thesection}{.5em}{\hspace{0in}}[\vspace{-.2\baselineskip}]
\titleformat{\subsection}[hang]{\bfseries}{\thesubsection}{.5em}{\hspace{0in}}[\vspace{-.2\baselineskip}]
\titleformat{\subsubsection}[hang]{\itshape}{\thesubsubsection}{.5em}{\hspace{0in}}[\vspace{-.2\baselineskip}]
\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1ex}{}{\vspace{-.2\baselineskip}}

%\titlespacing*{ ⟨command⟩}{⟨left⟩}{⟨beforesep⟩}{⟨aftersep⟩}[⟨right⟩]
\titlespacing{\section}{0pc}{1.5ex plus 0.1ex minus 0.2ex}{.5ex plus 0.1ex minus 0.1ex}
\titlespacing{\subsection}{0pc}{1.5ex plus 0.1ex minus 0.2ex}{.5ex plus 0.1ex minus 0.1ex}
\titlespacing{\subsubsection}{0pc}{1.5ex plus 0.1ex minus 0.2ex}{.5ex plus 0.1ex minus 0.1ex}

%%% Links and Table of Contents
\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks=TRUE,linkcolor=BrickRed,citecolor=MidnightBlue]{hyperref}

% \newif\ifanon
% \ifanon
% \footnote{We are awesome!}
% \else
% \fi
%% \abovedisplayshortskip=0pt
%% \belowdisplayshortskip=0pt
%% \abovedisplayskip=0pt
%% \belowdisplayskip=0pt
%% 
\begin{document}
%\VerbatimFootnotes

\begin{titlepage}
\maketitle
\thispagestyle{empty}

\begin{abstract} 
\citet{bowers2013sutva} showed that one could use Fisher's randomization-based
hypothesis testing framework to assess counterfactual causal models of
treatment propagation and spillover across social networks.  This research
note improves the statistical inference presented in \citet{bowers2013sutva}
(henceforth BFP) by substituting a test statistic based on a sum of squared
residuals and incorporating information about the fixed network for the simple
Kolmogorov-Smirnov test statistic  \citep[\S 5.4]{MylesHollander1999a}
they used. This note incrementally improves
the application of BFP's ``reasoning about interference'' approach. We do not
offer general results about test statistics for multi-parameter causal models on social
networks here, but instead hope to stimulate further,
and deeper, work on test statistics and sharp hypothesis testing.
\end{abstract}

Key Words: Causal effect; Interference; Randomized experiment;
Randomization inference; Fisher's Sharp Null Hypothesis; SUTVA

\end{titlepage}

%\input{titlepage.tex}
\setstretch{1.6} %1.6 or 1.37 which I think is "true double spacing"
%\onehalfspacing

\input{introduction.tex}

\cleardoublepage
%\appendix % is redefined by hyperref and a mess, do the appendix stuff by hand instead
% problems with pdf bookmarks and appendix caused by resetting section counter.
% solution here: http://www.latex-community.org/forum/viewtopic.php?f=45&t=4950
\begin{singlespacing}
%\renewcommand{\thesection}{Appendix \thesection}
%\setcounter{section}{0}
\newcounter{mycounter}
\setcounter{mycounter}{0}
%\renewcommand{\thesection}{Appendix \Alph{mycounter}}
\let\osection\section
\renewcommand{\section}{\stepcounter{mycounter}\osection}
\renewcommand\thesection{Appendix \Alph{mycounter}}
%\titleformat{\section}[hang]{\large\bfseries}{\thesection}{.5em}{\hspace{0in}}[]
%\input{codeexample}
%%\clearpage
%%\input{generalinterference}

\clearpage
\bibliographystyle{apsr}
\bibliography{../BIB/big}

% appendix 
% \clearpage
% \setlength{\parindent}{0pt}
% \addtolength{\parskip}{.5\baselineskip}
% \pagenumbering{arabic} %\setcounter{page}{1}
% \setcounter{mycounter}{0}
% \setcounter{secnumdepth}{0}
% %\let\osection\section
% %\renewcommand{\section}{\stepcounter{mycounter}\osection}
% \renewcommand\thesection{\arabic{mycounter}}
% \input{response}
%
\end{singlespacing}
%\bibliography{/Users/jwbowers/Documents/BIB/trunk/big}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End: