### 2016-01-04 response.md
Final response to editors for R1

Dear Professor Grimmer,

We really appreciate the constructive and timely feedback from you and the
reviewers. Here is our response to your questions and suggestions and those of
the reviewers.

# Overall

1) As Reviewer 1 suggests, how does the model perform as the assumptions
about spill overs deviate?  How sensitive are the inferences to the
particular assumptions you can make?

In the paper we clarified that we are focusing on testing sharp null
hypotheses arising from one model of how an experimental treatment might
propagate across a social network. We have added some discussion to the paper
highlighting the fact that the power of a test of a sharp null hypothesis
depends on *both* the model of causal effects *and* the test statistic (and
the effective sample size which itself depends on sample size and network
characteristics). We also note that, given randomization, any test statistic
ought to produce valid statistical inferences (i.e. tests that would reject
the truth no more than $100 \alpha \%$ of the time).

Since the point of the paper is to learn about test statistics, we do not systematically
explore alternative models of the propagation of causal effects across
networks. The original paper (BFP) on this method compared the performance of
tests when the data did and did not arise from the causal model and we point to
that paper for those who would like to pursue those set of questions. We did
add some discussion about the power of the SSR test statistic against different
kinds of simple causal effects and outcome distributions: the SSR test
statistic ought to be particularly sensitive to changes in mean differences
(after all, recall that the coefficient on a dichotomous variable in an least
squares model with a continuous outcome that minimizes the SSR is the
difference of means) and ought to be less sensitive to other differences than a
statistic like the KS test statistic.

2) Reviewer 2 asks you to carry forward your toy example with four units
through the entire analysis.  If this isn't possible, please provide the
simplest case that illuminate the core intuition of your method.

When we set about adding the four unit example to the different sections of the
paper, we realized that we were not enhancing intuition: the two test
statistics applied to an experiment of size 4 mainly reflected the fact that we
had very little data. Instead, we added a discussion of the SSR and KS test
statistics applied to a case with no interference and with very simple models
of causal effect. We also added a figure and discussion that tries to show a bit more
about how models of effects work in the context of sharp null hypothesis
testing. We hope that this helps with the intuition so that the use of the SSR
test statistic in an experiment on a social network with a complex model of
treatment propagation is easier to understand.

3) Reviewer 2 also asks you to provide intuition about when the KS-test
would outperform the Sum of Squared Residuals.  As part of a simulation or
analysis of a simple case, demonstrating when/why the KS test outperforms
will be useful for gaining intuition about why the SSR is a generally
superior method.

In the text we note that the SSR is particularly sensitive to differences in
means (or perhaps symmetric differences in tails that also produce a lot of
movement in the mean of one group relative to another group) while the KS-test
ought to have more power to detect asymmetric differences in the distributions
of the treated versus control groups --- for example, differences that change
variance without moving the means by stretching one tail of the distribution.

4) In letters we have an emphasis that new methods matter when applied to
real examples.  Can you demonstrate that a case where the SSR will change
the inferences we make when applied to a real world example?

We have added a section in which we replicate the analysis in Coppock's 2014
*Journal of Experimental Political Science* article "Information Spillovers:
Another Look at Experimental Estimates of Legislator Responsiveness".

# Response to Reviewer 1

We appreciate the positive review. In response to this review and the comments
from the editor and other reviewer, we now evaluate the SSR test statistic for
three non-interference models (the sharp null model, the constant additive
effects model, and the constant multiplicative effects model) and one new
interference model (the one used by Coppock in his *Journal of Experimental
Political Science* article).  We did not systematically search for other
spillover models from political science, but agree that such a survey of
models and now, test statistics, would be useful in the future.

# Response to Reviewer 2

We appreciate the positive review. We added a section in an attempt to provide
more intuition about test statistics, models of effects, and their
interaction. Continuing the 4 unit example did not, in the end, clarify
Fisher/Rosenbaum-style hypothesis testing as we would have hoped so we only
use that example in the beginning to illustrate the idea that each unit in a
study may have many more than two potential outcomes. We did add a section
showing a case where the KS test has more power than the SSR test -- in this
case we used a skewed outcome with no spillover rather than heterogeneous
exposure to spillover in order to keep things as simple as possible.